Secure Composition of Robust and Optimising Compilers

### 2023-08-05 tech/main.tex
add some rules
add more estep rules

\section{General Trace Model}\label{ap:def:gtm}

@@

>>>>

\section{General Trace Model}\label{ap:sec:gtm}

@@

\section{Programming Languages}\label{ap:sec:proglangs}

\subsection{$\Ltms$: Statically Typed, Temporal Memory Safe, Strict Cryptographic Constant Time}

$\Ltms$ trivially satisfies \gls*{scct}, because all variables/values are private ($\lock$) by default.
This is ,,enforced'' when projecting $\src{\Ltms}$ events to the general trace model presented in \Cref{ap:sec:gtm}.

\begin{gather*}
  \begin{aligned}
    \mi{(Expressions)}~&\src{\expr}\bnfdef\ \src{x} \mid \src{\valueexpr} \mid \src{\lbinop{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lget{x}{\expr}} \mid \src{\llet{x}{\lnew{\expr[_{1}]}{\expr[_{2}]}}{\expr[_{3}]}} \mid \src{\ldelete{x}} \mid \src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}} \\
                       & \mid \src{\lpair{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lproja{\expr}} \mid \src{\lprojb{\expr}} \mid \src{\llet{x}{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lreturn{\expr}} \mid \src{\lcall{g}{\expr}} \mid \src{\lifz{\expr}{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\labort}
  \end{aligned}
  \\
  \begin{aligned}
  \mi{(Types)}~&\src{\type}\bnfdef\src{\natt} \mid \src{\type[_{1}]} \times \src{\type[_{2}]}
  &
  \mi{(Functions)}~&\src{\function}\bnfdef\ \src{\lfunction{\textit{foo}}{x}{e}}
  &
\mi{(Libraries)}~&\src{\library}\bnfdef\ \src{\hole{\cdot}} \mid \src{\function},\src{\library}
  \end{aligned}\\
  \begin{aligned}
  \mi{(Component~Names)}~&\src{\commlib}\bnfdef\ \src{\hole{\cdot}} \mid \src{\text{foo}},\src{\commlib}
  &
  \mi{(Programs)}~&\src{\prog{\library_{\ctx}}{\library_{\comp}}}
  \end{aligned}
\end{gather*}

Variables are referred to as $x,y,z,a,b,c,\ldots$ while functions may be referred to as multi-character words, such as $\texttt{foo}$, as well as short-forms like $f,g,h$.
    \mi{(Control~Tags)}~&\src{\sandboxtag}\bnfdef\src{\ctx} \mid \src{comp}
    &
    \mi{(Communication~Tags)}~&\src{\comm}\bnfdef\ \src{\ctxtocomp} \mid\ \src{\comptoctx} \mid \src{\nocomm}
    &
    \mi{(Poison~Tags)}~&\src{\poison}\bnfdef \src{\poisoned} \mid \src{\poisonless}
  \end{aligned}\\
    \mi{(Continuation~Stacks)}~&\src{\ectxstack}\bnfdef\src{\hole{\cdot}} \mid \src{\tup{K}{g}},\src{\ectxstack}
    &
    \mi{(Control~Flow~States)}~&\src{\cfstate}\bnfdef\src{\tup{\library}{\ectxstack}}
  	\\
    \mi{(Locations)}~&\src{\loc}\in\nat
    &
    \mi{(Substitutions)}~&\src{\substvar}\bnfdef \src{\hole{\cdot}} \mid \src{\subst{\valueexpr}{x}},\src{\gamma}
  	\\
    \mi{(Stores)}~&\src{\Delta}\bnfdef\src{\hole{\cdot}} \mid \src{x\mapsto\tup{\loc}{\sandboxtag,\poison,n}},\src{\Delta}
    &
    \mi{(Heaps)}~&\src{H}\bnfdef\src{\hole{\cdot}} | \src{\valueexpr},\src{H}
  \end{aligned}\\
    \mi{(Memory~States)}~&\src{\memstate}\bnfdef\src{\tup{H^{\ctx}}{H^{\comp};\Delta}}
    &
    \mi{(States)}~&\src{\statevar}\bnfdef\src{\tup{\cfstate}{\sandboxtag,\memstate}}
    &
    \mi{(Runtime~Terms)}~&\src{\runtimetermvar}\bnfdef\src{\rtt{\statevar}{\expr}} \mid \src{\lightning}
Throughout the paper, the substitution notation is also used to update entries in states $\statevar$.
    \mi{(Pre\text{-}Events)}~\src{\event[_{b}]}&\bnfdef \src{\ev{Alloc}\ \loc\ \valueexpr} \mid \src{\ev{Dealloc}\ \loc} \mid \src{\ev{Get}\ \loc\ \valueexpr} \mid \src{\ev{Set}\ \loc\ \valueexpr\ \valueexpr[']} \mid \src{\ev{Call}\ \comm\ g\ \valueexpr} \mid \src{\ev{Ret}\ \comm\ \valueexpr} \mid \src{\ev{Start}} \mid \src{\ev{End}\ \valueexpr}\\
  \end{aligned}\\
    \mi{(Events)}~\src{\event}&\bnfdef \src{\emptyevent} \mid \src{\lightning} \mid \src{\tup{\event[_{b}]}{\sandboxtag}} \\
  \judgbox{\pstepto{\src{\runtimetermvar}}{\src{\runtimetermvar[']}}{\src{\event}}}{,,$\src{\runtimetermvar}$ does one primitive step to $\src{\runtimetermvar[']}$ emitting event $\src{\event}$.''}
  \typerule{$\src{\Ltms}-e-\text{binop}$}{
    \src{\lbinop{\expr[_1]}{\expr[_2]}}=\src{\expr[_3]}
    \pstepto{\src{\rtt{\statevar}{\lbinop{\expr[_1]}{\expr[_2]}}}}{\src{\rtt{\statevar}{\expr[_3]}}}{\src{\emptyevent}}
  }{ap:ltms-e-binop}
  \typerule{$\src{\Ltms}-e-\text{get}-\in$}{
    \src{\sandboxtag}=\src{\statevar.\sandboxtag} &
    \src{\statevar.\store(x)}=\src{\tup{\loc}{\sandboxtag,\poison,m}} \\
    \src{\loc}+\src{n}\in\operatorname{dom}\src{\statevar.\heap[^{\sandboxtag}](\loc+n)}
    \pstepto{\src{\rtt{\statevar}{\lget{x}{n}}}}{\src{\rtt{\statevar}{\statevar.\heap[^{\sandboxtag}](\loc+n)}}}{\src{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}}
  }{ap:ltms-e-get-in}
  \typerule{$\src{\Ltms}-e-\text{get}-\notin$}{
    \src{\sandboxtag}=\src{\statevar.\sandboxtag} &
    \src{\statevar.\store(x)}=\src{\tup{\loc}{\sandboxtag,\poison,m}} \\
    \src{\loc}+\src{n}\notin\operatorname{dom}\src{\statevar.\heap[^{\sandboxtag}](\loc+n)}
    \pstepto{\src{\rtt{\statevar}{\lget{x}{n}}}}{\src{\rtt{\statevar}{1729}}}{\src{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}}
  }{ap:ltms-e-get-notin}
  \typerule{$\src{\Ltms}-e-\text{set}-\in$}{
    \src{\sandboxtag}=\src{\statevar.\sandboxtag} &
    \src{\statevar.\store(x)}=\src{\tup{\loc}{\sandboxtag,\poison,m}} \\
    \src{\loc}+\src{n}\in\operatorname{dom}\src{\statevar.\heap[^{\sandboxtag}]} &
    \src{\heap[']}=\seth{\src{\statevar.\heap[^{\sandboxtag}]}}{\src{\loc}+\src{n}}{\src{\valueexpr}}
    \pstepto{\src{\rtt{\statevar}{\lset{x}{n}{\valueexpr}}}}{\src{\rtt{\statevar\subst{\heap[^{\sandboxtag}]}{\heap[']}}{\valueexpr}}}{\src{\tup{\ev{Set}\ \loc\ n\ \valueexpr}{\sandboxtag}}}
  }{ap:ltms-e-set-in}
  \typerule{$\src{\Ltms}-e-\text{set}-\notin$}{
    \src{\sandboxtag}=\src{\statevar.\sandboxtag} &
    \src{\statevar.\store(x)}=\src{\tup{\loc}{\sandboxtag,\poison,m}} \\
    \src{\loc}+\src{n}\notin\operatorname{dom}\src{\statevar.\heap[^{\sandboxtag}]}
    \pstepto{\src{\rtt{\statevar}{\lset{x}{n}{\valueexpr}}}}{\src{\rtt{\statevar}{\valueexpr}}}{\src{\tup{\ev{Set}\ \loc\ n\ \valueexpr}{\sandboxtag}}}
  }{ap:ltms-e-set-notin}
  \typerule{$\src{\Ltms}-e-\text{new}$}{
    \fresh{\src{\statevar}}{\src{z}} &
    \fresh{\src{\statevar}}{\src{\loc}} &
    \src{\heap[^{\sandboxtag}_1]}=\growh{\src{\statevar.\heap[^{\sandboxtag}]}}{\src{n}}{\src{\valueexpr}} &
    \src{\store[_1]}=\src{\storeel{z}{\loc}{\statevar.\sandboxtag}{\poisonless}{n},\statevar.\store}
    \pstepto{\src{\rtt{\statevar}{\lnew{x}{\valueexpr}{n}{\expr}}}}{\src{\rtt{\statevar\subst{\heap[^{\sandboxtag}]}{\heap[^{\sandboxtag}_1]}\subst{\store}{\store[_1]}}{\expr\subst{z}{x}}}}{\src{\tup{\ev{Alloc}\ \loc\ n}{\statevar.\sandboxtag}}}
  }{ap:ltms-e-new}
  \typerule{$\src{\Ltms}-e-\text{dealloc}$}{
    \src{\statevar.\store(x)} = \src{\tup{\loc}{\statevar.\sandboxtag,\poison,n}} &
    \src{\store[_1]}=\seth{\src{\statevar.\store}}{\src{x}}{\src{\tup{\loc}{\statevar.\sandboxtag,\poisoned,n}}}
    \pstepto{\src{\rtt{\statevar}{\ldelete{x}}}}{\src{\rtt{\statevar\subst{\store}{\store[_1]}}{0}}}{\src{\tup{\ev{Dealloc}\ \loc}{\statevar.\sandboxtag}}}
  }{ap:ltms-e-dealloc}
  \typerule{$\src{\Ltms}-e-\text{let}$}{
    \pstepto{\src{\rtt{\statevar}{\llet{x}{\valueexpr}{\expr}}}}{\src{\rtt{\statevar}{\expr\subst{x}{\valueexpr}}}}{\src{\emptyevent}}
  }{ap:ltms-e-let}
  \typerule{$\src{\Ltms}-e-\text{ifz}-yes$}{
    \pstepto{\src{\rtt{\statevar}{\lifz{0}{\expr[_1]}{\expr[_2]}}}}{\src{\rtt{\statevar}{\expr[_1]}}}{\src{\emptyevent}}
  }{ap:ltms-e-ifz-yes}
  \typerule{$\src{\Ltms}-e-\text{ifz}-no$}{
    \src{n}\not=\src{0}
    \pstepto{\src{\rtt{\statevar}{\lifz{n}{\expr[_1]}{\expr[_2]}}}}{\src{\rtt{\statevar}{\expr[_2]}}}{\src{\emptyevent}}
  }{ap:ltms-e-ifz-no}
  \typerule{$\src{\Ltms}-e-\text{proj}-a$}{
    \pstepto{\src{\rtt{\statevar}{\lproja{\lpair{\valueexpr[_1]}{\valueexpr[_2]}}}}}{\src{\rtt{\statevar}{\valueexpr[_1]}}}{\src{\emptyevent}}
  }{ap:ltms-e-proj-a}
  \typerule{$\src{\Ltms}-e-\text{proj}-b$}{
    \pstepto{\src{\rtt{\statevar}{\lprojb{\lpair{\valueexpr[_1]}{\valueexpr[_2]}}}}}{\src{\rtt{\statevar}{\valueexpr[_2]}}}{\src{\emptyevent}}
  }{ap:ltms-e-proj-a}
  \typerule{$\src{\Ltms}-e-\text{abort}$}{
    \pstepto{\src{\rtt{\statevar}{\labort}}}{\src{\lightning}}{\src{\lightning}}
  }{ap:ltms-e-abort}
  \judgbox{\src{\commlib}\vdash\src{foo}:\src{\sandboxtag}\dashv\src{\comm},\src{\sandboxtag'}}{,,Determine communication tag $\src{\comm}$ and control tag $\src{\sandboxtag'}$ by checking the current control $\src{\sandboxtag}$,\\the called function $\src{foo}$, and list of component names $\src{\commlib}$.''}
  \typerule{$\src{\Ltms}-\text{switch}-\ctx-\text{to}-\comp$}{
    \src{foo}\in\src{\commlib}
    \src{\commlib}\vdash\src{foo}:\src{\ctx}\dashv\src{\ctxtocomp},\src{\comp}
  }{ap:ltms-switch-ctx-to-comp}
  \typerule{$\src{\Ltms}-\text{switch}-\comp-\text{to}-\ctx$}{
    \src{foo}\notin\src{\commlib}
    \src{\commlib}\vdash\src{foo}:\src{\comp}\dashv\src{\comptoctx},\src{\ctx}
  }{ap:ltms-switch-comp-to-ctx}
  \typerule{$\src{\Ltms}-\text{noswitch}-\ctx$}{
    \src{foo}\notin\src{\commlib}
    \src{\commlib}\vdash\src{foo}:\src{\ctx}\dashv\src{\nocomm},\src{\ctx}
  }{ap:ltms-noswitch-ctx}
  \typerule{$\src{\Ltms}-\text{noswitch}-\comp$}{
    \src{foo}\in\src{\commlib}
    \src{\commlib}\vdash\src{foo}:\src{\comp}\dashv\src{\nocomm},\src{\comp}
  }{ap:ltms-noswitch-comp}
  \judgbox{\estepto[]{\src{\runtimetermvar}}{\src{\runtimetermvar[']}}{\src{\event}}}{,,$\src{\runtimetermvar}$ does one contextual step to $\src{\runtimetermvar[']}$ emitting event $\src{\event}$.''}
  \typerule{$\src{\Ltms}-e-\text{ectx}$}{
    \pstepto{\src{\rtt{\statevar}{\expr}}}{\src{\rtt{\statevar[']}{\expr[']}}}{\src{\event}}
    \estepto[]{\src{\rtt{\statevar}{\ectx\hole{\expr}}}}{\src{\rtt{\statevar[']}{\ectx\hole{\expr[']}}}}{\src{\event}}
  }{ap:ltms-e-ectx}
  \typerule{$\src{\Ltms}-e-\text{stuck}$}{
    \pstepto{\src{\rtt{\statevar}{\expr}}}{\src{\lightning}}{\src{\lightning}}
    \estepto[]{\src{\rtt{\statevar}{\ectx\hole{\expr}}}}{\src{\lightning}}{\src{\lightning}}
  }{ap:ltms-e-stuck}
  \typerule{$\src{\Ltms}-e-\text{start}$}{
    \src{\lfunction{\text{main}}{x}{e}} \in \src{\statevar.\library} &
    \src{\statevar.\ectxstack}=\src{\hole{\cdot}}
    \estepto[]{\src{\rtt{\statevar}{\lcall{\text{main}}{0}}}}{\src{\rtt{\statevar\subst{\ctx}{\sandboxtag}\subst{\tup{\hole{\cdot}}{\text{main}},\hole{\cdot}}{\ectxstack}}{\expr\subst{0}{x}}}}{\src{\tup{\ev{Start}}{\comp}}}
  }{ap:ltms-e-start}
  \typerule{$\src{\Ltms}-e-\text{ret}$}{
    \src{\statevar.\ectxstack}=\src{\tup{\ectx}{\text{foo}}},\src{\ectxstack[']} &
    \src{\commlib}\vdash\src{\text{foo}}:\src{\statevar.\sandboxtag}\dashv\src{\comm}
    \estepto[]{\src{\rtt{\statevar}{\ectx[']\hole{\lreturn{\valueexpr}}}}}{\src{\rtt{\statevar\subst{\neg\statevar.\sandboxtag}{\sandboxtag}\subst{\ectxstack[']}{\ectxstack}}{\ectx\hole{\valueexpr}}}}{\src{\tup{\ev{Ret}\ \comm\ \valueexpr}{\statevar.\sandboxtag}}}
  }{ap:ltms-e-ret}
  \typerule{$\src{\Ltms}-e-\text{call}$}{
    \src{\lfunction{\text{foo}}{x}{e}} \in \src{\statevar.\library} &
    \src{\commlib}\vdash\src{\text{foo}}:\src{\statevar.\sandboxtag}\dashv\src{\comm}
    \estepto[]{\src{\rtt{\statevar}{\ectx\hole{\lcall{\text{foo}}{\valueexpr}}}}}{\src{\rtt{\statevar\subst{\neg\statevar.\sandboxtag}{\sandboxtag}\subst{\tup{\ectx}{\text{foo}},\statevar.\ectxstack}{\ectxstack}}{\expr\subst{\valueexpr}{x}}}}{\src{\tup{\ev{Call}\ \comm\ \text{foo}\ \valueexpr}{\statevar.\sandboxtag}}}
  }{ap:ltms-e-call}
  \typerule{$\src{\Ltms}-e-\text{end}$}{
    \src{\statevar.\ectxstack}=\src{(\hole{\cdot},main),\hole{\cdot}}
    \estepto[]{\src{\rtt{\statevar}{\ectx\hole{\lreturn{\valueexpr}}}}}{\src{\rtt{\statevar\subst{\comp}{\sandboxtag}\subst{\tup{\hole{\cdot}}{\text{main}},\hole{\cdot}}{\ectxstack}}{\valueexpr}}}{\src{\tup{\ev{End}\ \valueexpr}{\ctx}}}
  }{ap:ltms-e-end}
